The human in the loop: volunteer-based metacomputers as a socio-technical system

### 2016-04-24 Alife16/paper.tex
Shortening the state of the art to cut the paper down to 8 pages

On the other hand, we are also interested in measuring the performance
of volunteer computing systems, an area in which there have been
relatively few efforts.
There were some initial attempts to avoid the differences in performance
that could be obtained from volunteers  by making
the algorithm adaptive to the kind of resources allotted to it
\citep{milani2004online}, which is actually not such a big problem in
algorithms such as the EA that can be easily 
parallelized via population splitting or by farming out the evaluation to all
the nodes available. Lately, several approaches have focused on the
fault-tolerance of volunteer systems
\citep{gonzalez2010characterizing} which it can, of course, be studied in
the more general context of distributed computing 
\citep{nogueras2015studying}  or included in a more general 
study of the performance of the EA itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. This performance cannot be
measured without first understanding the dynamics of this kinds of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. On the contrary, in this paper, we will use {\em
  actual} volunteers, anonymous persons that are requested to join an
experiment through social networks.

>>>>

The number of users is key in the performance of these systems, but it
also essential to adapt
the algorithm itself to the available resources, as shown by
\citep{milani2004online}, although EAs can be readily distributed via
population splitting or by farming out the evaluation to all 
the nodes available. However, user churn affects experiment performance 
\citep{gonzalez2010characterizing,nogueras2015studying} and also the
performance of the algorithm itself
\citep{DBLP:journals/gpem/LaredoBGVAGF14}. All these issues imply that
a the performance of a volunteer system cannot be
measured without first understanding its dynamics. Initial
work was done for peer to peer systems by Stutzbach et
al. \citep{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \citep{churn08,laredo2008rcp}. A similar
study was performed by Martínez et al. on the Capataz system
\citep{martinez2015capataz}; however, in this case the number of
computers used was known in advance and the main focus was on
measuring the speed up and how job bundling helped to reduce overhead and
enhance performance. On the contrary, in this paper, we will use {\em
  actual} volunteers.
